Using Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2020-12-09 thesis/thesis.tex
thesis: Update title
thesis: Update Introduction chapter

%---------------------------
% Título del trabajo y autor
%---------------------------
\title{Open Clusters Characterization in Gaia DR2 Using ML Algorithms}
\author{Carlos David Álvaro Yunta}
\date{12th of November, 2020}
\director{César Augusto Guzmán Álvarez}
\nombreciudad{Madrid, Spain}

@@

% TODO: 1.1 Motivation

% ¿Cuál es el problema a tratar?
% ¿Cuáles crees que son las causas?
% ¿Por qué es relevante el problema?

% TODO: 1.2 Planteamiento

% ¿Cómo se podría solucionar el problema?
% ¿Qué es lo que se propone?
% Descripción del objetivo del trabajo en términos generales

% TODO: 1.3 Estructura de la memoria

% Descripción breve de lo que se va a contar en cada uno de los capítulos siguientes.

@@

Although most stars in the Milky Way are presented isolated, it is considered that most of them (or even all)
are formed in clustered environments and spend a period of time gravitationally bound with their siblings embedded
in their original molecular cloud
\cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
Galactic tidal forces and mechanisms that involves the gas loss driven by stellar feedback are other causes of disruption
\cite{brinkmann2017bound}.
Nevertheless, a small fraction of these systems will survive in the initial state and persist bound in bigger timescales.

Young OCs allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
On the other side, older OCs give us information about stellar processes and how the galactic disk evolves.
Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy
\cite{cantat2016abundances}.

@@

Stars that belong to a same OC share relative positions, inherited from their original gas cloud.
So their distance from the Earth is similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
They also share similar values of proper motion, both in right ascension and declination. Another property they share is their chemical composition.
Thus, their metallicity must be uniform, since these stars were born from the same gas cloud and at the same time stage.
However, to take this last property into account, we are faced with the drawback that this parameter is poorly reported in Gaia DR2 database.

@@

We will take advantage of these properties only to validate our characterization of OCs, but not to determine what stars belong to them. As we will explain later,
we will only use dynamic properties such as proper motions and parallax to characterize open clusters.

Although the members of a cluster appear in the observation visual field as an overdensity in positions, as shown in Figure \ref{fig:pos_ngc_2682}, these coordinates
are not useful to separate those stars that belong to the cluster from the other which do not. However, if we look for overdensities in the proper motions configuration space,
it is possible, at least at first instance, to assume a possible membership delimitation. See Figure \ref{fig:pm_ngc_2682}.

@@

@@

Until now, we have focus our study in Melotte 22 OC. Since our main aim is to get a model that fits well for a wide range of clusters,
we present now the results obtained for a selection of clusters with different topologies.

>>>>

%---------------------------
% Título del trabajo y autor
%---------------------------
\title{Using Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
\author{Carlos David Álvaro Yunta}
\date{12th of November, 2020}
\director{César Augusto Guzmán Álvarez}
\nombreciudad{Madrid, Spain}

@@

@@

Although most stars in the Milky Way are presented isolated, it is considered that most of them (or even all)
are formed in clustered environments and spend a period of time gravitationally bound with their siblings embedded
in their original molecular cloud \cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
Galactic tidal forces and mechanisms that involves gas loss driven by stellar feedback are other causes of disruption
\cite{brinkmann2017bound}.
Nevertheless, a small fraction of these systems will survive in the initial state and persist bound in bigger timescales.

Young OCs allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
On the other hand, older open clusters give us information about stellar processes and how the galactic disk evolves.
Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy
\cite{cantat2016abundances}.

@@

Stars that belong to a same OC share relative positions, inherited from their original gas cloud.
So their distance to the Earth is similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
They also share similar values of proper motion, both in right ascension and declination.
Another property they share is their chemical composition.
Thus, their metallicity must be uniform, since these stars were born from the same gas cloud and at the same time stage.
However, to take this last property into account, we are faced with the drawback that this parameter is poorly reported in Gaia DR2 database.

@@

We will take advantage of these properties only to validate our characterization of OCs,
but not to determine what stars belong to them. As we will explain later,
we will only use dynamic properties such as proper motion and parallax to characterize open clusters.

Although the members of a cluster appear in the observation visual field as an overdensity in positions,
as shown in Figure \ref{fig:pos_ngc_2682},
these coordinates are not useful to separate those stars that belong to the cluster from the other which do not.
However, if we look for overdensities in the proper motions configuration space, it is possible, at least at first instance,
to assume a possible membership delimitation (see Figure \ref{fig:pm_ngc_2682}.

@@

We start enumerating the aims we try to achieve with this work.
Then, we expose the current \emph{state of the art}, mentioning the \emph{Gaia Mission} used as data source in this work and
describing other related works that try to solve this problem.
As a last point of the state of the art, we make a brief summary of some of the today available clustering algorithms.
After the state of the art, we explain our method to improve open cluster characterization.
Next, we show an alternative method based on current procedures. We have used this alternative method as a validation method to test ours.
Then, a comparison between a set of results obtained with our method and the results obtained for the same clusters with the validation method is presented.
Finally, we extract some conclusions based on the obtained results and present further researches that can be done in light of the results.

@@

Until now, we have focused our study in Melotte 22 OC. Since our main aim is to get a model that fits well for a wide range of clusters,
we present now the results obtained for a selection of clusters with different topologies.
